### 2014-09-06 overview.tex

\subsection{Preliminaries}
We first introduce some notation used throughout this paper. Intuitively, a 
trace is a sequence of events recorded during the observed run. An 
event, $e=<t, id, inst>$, is a concrete representation of the runtime 
execution of a static instruction $inst$, such that (i) $t$ refers to the 
thread; (ii) $id$ is a unique identifier of the event (unless specified 
otherwise, we use the index of an event in the trace as its id); 
and (iii) $inst$ is a static three-address instruction (involving at most three operands).

@@

%TODO this section is inconsistent with others. change it!
\subsection{Constraint System}
In this section, we are going to informally explain the two relaxations
we make which allow us to catch the two races missed by existing 
predictive analyses.
Our analysis takes the program and a trace as the input.  
Suppose it starts with the following trace of the program in  
Figure \ref{fig:running}:  $e_1 e_2 e_3 e_4 e_5$, where the line number 
is used as the event id (subscript).

\smallskip
\noindent
{\em I. Value Relaxation\ } Our first relaxation is to allow variables
to have different values as long as the same thread local path is preserved.
It does not require data dependences or variable values as in~\cite{...}.
This allows us to catch the racy pair $(e_2, e_5)$.
%A potential race is between 
%the pair, $(e_2, e_5)$, as they access the shared variable $x$ from 
%different threads and one is a write. 
Given a trace, $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \},
 O, \theta>$, our analysis computes a new trace,  
$\tau'=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O', \theta'>$.
 $\tau'$ preserves the same event sequence for each thread, i.e., $T_1$ still 
executes $e_1 e_2 e_3$ and $T_2$ still executes $e_4 e_5$, but it reschedules 
the events from different threads so that the potential race pair runs 
concurrently, i.e., $e_2$ and $e_5$ run concurrently. When the schedule 
changes, the variables may get different values. Therefore, we need to 
compute both the new schedule $O'$ and the new value mapping $\theta'$, in 
order to witness a race. To guarantee that $\tau'$ is a feasible trace of 
the program $P$, the computation needs to respect a set of constraints, 
which include the value constraints, control flow constraints and 
synchronization constraints~\footnote{We omit the synchronization constraints 
in this paper, which are already well explained in existing 
techniques~\cite{yannis, pecan}.}. We explain the constraints 
briefly in the following. Details are in Section~\ref{sec:relax1} 
and the soundness guarantee is explained in Section~\ref{sec:guarantee}.

@@

\smallskip
\noindent
{\em II. Path Relaxation\ } 
%Next, we show how we catch the second missing race
%($e_2$, $e_7$) with another relaxation. 
Besides the first relaxation, which focuses on the given trace, the second
relaxation is to even reason about unexecuted branches as long as they
can be precisely modeled. In our 
running example, this is essential to discovering the race between 
lines {\tt 2} and {\tt 7}. We explore the unexecuted branch statically 
and collect constraints during the exploration. We then form a disjunction  
of the original constraints and the collected constraints to include more 
possible  paths. The solver may pick up the unexplored branch to detect 
a race. Back to our example, event $e_7$ is in the unexplored branch, \tool
is able to identify the race between $e_7$ and $e_2$ after encoding 
the unexplored branch.

@@

\smallskip
\noindent
{\em Highlights \ } \tool finds two more pairs of races 
beyond existing analyses~\cite{yannis, jeff}: $(e_2, e_5)$ and $(e_2, e_7)$.
For the first race, we do not require $e_4$ to read from $e_3$ to 
retain the original value as in the
existing approaches. Rather, we allow $e_4$ to read from $e_1$ while 
still preserving the thread local path. This relaxation allows 
the events after $e_4$ to run concurrently with the events before 
$e_3$, leading to the detection of the first race.  For the second race, 
we cover unexecuted branches and query the solver to compute a feasible 
schedule that exercises them. 
%The neighboring branches are likely to 
%contain racy events if the executed branches contain racy events.

>>>>

\subsection{Preliminaries}
We first introduce some notation used throughout this paper. Intuitively, a 
trace is a sequence of events recorded during the observed run. An 
event, $e=<t, id, inst>$, is a concrete representation of the runtime 
execution of a static instruction $inst$, such that (i) $t$ refers to the 
thread; (ii) $id$ is a {\em unique} identifier of the event (unless specified 
otherwise, we use the index of an event in the trace as its id); 
and (iii) $inst$ is a static three-address instruction (involving at most three operands).

@@

\subsection{Constraint System}
In this section, we are going to informally explain the two relaxations
we make which allow us to catch the two races missed by existing 
predictive analyses.
Our analysis takes the program and a trace as the input.  
Suppose it starts with the following trace of the program in  
Figure \ref{fig:running}:  $e_1 e_2 e_3 e_4 e_5$, where the line number 
is used as the event id (subscript).

\smallskip
\noindent
{\em I. Value Relaxation\ } Our first relaxation is to allow variables
to have different values as long as the same thread local path is preserved.
It does not require data dependences or same variable values as in~\cite{yannis, pldi14, Said:2011}.
This allows us to catch the racy pair $(e_2, e_5)$.
%A potential race is between 
%the pair, $(e_2, e_5)$, as they access the shared variable $x$ from 
%different threads and one is a write. 
Given a trace, $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \},
 O, \theta>$, our analysis computes a new trace,  
$\tau'=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O', \theta'>$.
 $\tau'$ preserves the same event sequence for each thread, i.e., $T_1$ still 
executes $e_1 e_2 e_3$ and $T_2$ still executes $e_4 e_5$, but it reschedules 
the events from different threads so that the potential race pair runs 
concurrently, i.e., $e_2$ and $e_5$ run concurrently. When the schedule 
changes, the variables may get different values. Therefore, we need to 
compute both the new schedule $O'$ and the new value mapping $\theta'$, in 
order to witness a race. To guarantee that $\tau'$ is a feasible trace of 
the program $P$, the computation needs to respect a set of constraints, 
which include the value constraints, control flow constraints and 
synchronization constraints~\footnote{We omit the synchronization constraints 
in this paper, which are already well explained in existing 
techniques~\cite{yannis, pecan}.}. We explain the constraints 
briefly in the following. Details are in Section~\ref{sec:relax1} 
and the soundness guarantee is explained in Section~\ref{sec:guarantee}.

@@

\smallskip
\noindent
{\em II. Path Relaxation\ } 
%Next, we show how we catch the second missing race
%($e_2$, $e_7$) with another relaxation. 
Besides the first relaxation, which focuses on the given trace, the second
relaxation is to even reason about unexecuted branches as long as they
can be precisely modeled. In our 
running example, this is essential to discovering the race between 
lines {\tt 2} and {\tt 7}. We explore the unexecuted branch statically 
and collect constraints during the exploration. We then form a disjunction  
of the original constraints and the collected constraints to include more 
possible  paths. The solver may pick up the unexplored branch to detect 
a race. Back to our example, event $e_7$ is in the unexplored branch, \tool\ 
is able to identify the race between $e_7$ and $e_2$ after encoding 
the unexplored branch.

@@

\smallskip
\noindent
{\em Highlights \ } \tool\  finds two more pairs of races 
beyond existing analyses~\cite{yannis, pldi14}: $(e_2, e_5)$ and $(e_2, e_7)$.
For the first race, we do not require $e_4$ to read from $e_3$ to 
retain the original value as in the
existing approaches. Rather, we allow $e_4$ to read from $e_1$ while 
still preserving the thread local path. This relaxation allows 
the events after $e_4$ to run concurrently with the events before 
$e_3$, leading to the detection of the first race.  For the second race, 
we cover unexecuted branches and query the solver to compute a feasible 
schedule that exercises them. 
%The neighboring branches are likely to 
%contain racy events if the executed branches contain racy events.
